# Bike sharing
The model tries to predict the total count of bikes rented on a new test set

**Importing dataset from "https://www.kaggle.com/c/bike-sharing-demand/data"**
```{r}
# Loading required libraries
library(dplyr)
library(lubridate)
library(ggplot2)
library(reshape2)
library(ranger)
train=read.csv("../input/train.csv")
test=read.csv("../input/test.csv")

# Visualizing dataset
str(train);str(test)
any(is.na(train));any(is.na(test))   # Checking for NA's
all(train$casual + train$registered == train$count)

# Converting non-numerical features to categorical variables
factor_features=c('season','holiday','workingday','weather')
train[,factor_features]=lapply(train[,factor_features],as.factor)
test[,factor_features]=lapply(test[,factor_features],as.factor)
```
**Checking for correlation between numeric features**
```{r}
numeric_features=c('temp','atemp','humidity','windspeed','casual','registered','count')
cormat=cor(train[,numeric_features])
melted_cormat=melt(cormat)
head(melted_cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + ggtitle("Heatmap for correlation matrix of the dataset")
                                                    ### can use geom_raster() instead for many features 
```

**Extracting all important features from datetime**
```{r}
train_date=ymd_hms(train$datetime);             test_date=ymd_hms(test$datetime)   # yearmonthday_hourmonthsecond()
                                                                                   # the given datetime format
train$year=factor(year(train_date));            test$year=factor(year(test_date))
train$month=factor(month(train_date));          test$month=factor(month(test_date))
train$day=factor(day(train_date));              test$day=factor(day(test_date))
train$hour=factor(hour(train_date));            test$hour=factor(hour(test_date))
train$Weekday=factor(wday(train_date,label=T)); test$Weekday=factor(wday(test_date,label=T))

day_levels=append(levels(train$day),levels(test$day))                              # Making levels similar for day
levels(train$day)=day_levels
levels(test$day)=day_levels
```

**Visualizing hourly trends of count grouped by Season and Weekdays**
```{r}
season_summary = train %>% group_by(season,hour) %>% summarise(count=mean(count))
weekday_summary= train %>% group_by(Weekday,hour) %>% summarise(count=mean(count))

ggplot(data=season_summary,aes(x=hour,y=count,group=season,colour=season)) + geom_point() + geom_line()
ggplot(data=weekday_summary,aes(x=hour,y=count,group=Weekday,colour=Weekday)) + geom_point() + geom_line()
```
**Above plots give an idea of the trend followed on weekends(i.e. Sat & Sun) vs workingdays                                                 
It is also clear that there are few peakhours for both weekdays and weekends**
```{r}
# Creating new feature for weekends
train$weekend = factor(as.integer(train$Weekday %in% c('Sun','Sat')))
test$weekend = factor(as.integer(test$Weekday %in% c('Sun','Sat')))

# New feature for peakhours
train$peakhours = factor(as.integer((train$weekend==1 & train$hour %in% c(11,12,13,14,15,16,17)) | (train$weekend==0 & train$hour %in% c(8,17,18,19))))
test$peakhours = factor(as.integer((test$weekend==1 & test$hour %in% c(11,12,13,14,15,16,17)) | (test$weekend==0 & test$hour %in% c(8,17,18,19))))
```

**Creating scatter plots matrix for some insight**
```{r}
pairs(formula=~hour+day+weekend+weather+atemp+humidity+windspeed+peakhours+count,data=train)
```

**Finally creating a model using ranger (Boosted version of RandomForest)**
```{r}
train$count=log1p(train$count)                     # Scaling to log(1+x)
train$casual=log1p(train$casual)
train$registered=log1p(train$registered)

# Model on count
model = ranger(formula = count~season+holiday+workingday+weather+temp+atemp+humidity+windspeed+year+month+day+hour+Weekday+weekend+peakhours, data=train)
prediction=predict(model,test)

# Model on casual
model1 = ranger(formula = casual~season+holiday+workingday+weather+temp+atemp+humidity+windspeed+year+month+day+hour+Weekday+weekend+peakhours, data=train)
prediction1=predict(model1,test)

# Model on registered
model2 = ranger(formula = registered~season+holiday+workingday+weather+temp+atemp+humidity+windspeed+year+month+day+hour+Weekday+weekend+peakhours, data=train)
prediction2=predict(model2,test)

final_prediction=(expm1(prediction$predictions)+expm1(prediction1$predictions)+expm1(prediction2$predictions))/2
                                              # As casual+registered=count
submit = data.frame(datetime=test$datetime,count=final_prediction)
head(submit)
write.csv(submit,"submit.csv",row.names=F)
```

